
```{r}
library(rstudioapi)
library(dplyr)
library(cluster)
library(ggplot2)
library(lubridate)
```

```{r}
set.seed(7004)
main_dir<- dirname(dirname(rstudioapi::getSourceEditorContext()$path))
datadir<- paste0(main_dir,"/data")
clus_data_path<- paste0(datadir,"/data_c.csv")

clus_data<- read.csv(clus_data_path)
```


Besides time series modelling, we also do a clustering to identify cities (location) with similar air quality. We cluster them by using only the latest 1 year (2015-01-01 to 2015-12-31) of data, with the mean value of numerical variables (SO_2, NO_2, RSPM, SPM, PM_2.5) as the features.
```{r}
start_date<- ymd("2015-01-01")
rec_clus_data<- clus_data[clus_data$date > start_date, ]

### remove rows with missing location
rec_clus_data<- rec_clus_data[!is.na(rec_clus_data$location), ]

loc_data<- rec_clus_data %>% group_by(location) %>%
  summarise(mean_so2 = mean(so2, na.rm=TRUE),
            mean_no2 = mean(no2, na.rm=TRUE),
            mean_rspm = mean(rspm, na.rm=TRUE),
            mean_spm = mean(spm, na.rm=TRUE),
            mean_pm2_5 = mean(pm2_5, na.rm=TRUE)
  ) %>% data.frame()

head(loc_data)
```

Find and impute missing values
```{r}
num_loc = nrow(loc_data)

for (col in colnames(loc_data) ){
  num_na<- sum(is.na(loc_data[col]))
  perc_na<- round(100*num_na/num_loc, digits=2)
  print(paste(col, num_na, perc_na))
}
```
We see that SPM (100% missing) and PM2.5 (75.2%) have very large amount of missing values. We can only drop these features as imputing them will introduce large bias in the clustering model. 
For SO_2 and NO_2, we will impute using mean value.
```{r}
loc_data$mean_so2[is.na(loc_data$mean_so2)]= mean(loc_data$mean_so2, na.rm = TRUE)
loc_data$mean_no2[is.na(loc_data$mean_no2)]= mean(loc_data$mean_no2, na.rm = TRUE)

#### set location as row names
rownames(loc_data)<- loc_data$location

#### drop unnecessary columns
clean_loc_data<- loc_data %>% select(-c(mean_spm, mean_pm2_5, location))

head(clean_loc_data)
```
Next, we scale the features so that their magnitude are of the same order.
```{r}
scaled_loc_data<- clean_loc_data
for (mean_col in c("mean_so2", "mean_no2", "mean_rspm")){
 scaled_loc_data[mean_col]<- scale(scaled_loc_data[mean_col], center=TRUE, scale=TRUE)
}
```

Now we can do clustering. We use the elbow method to obtain a suitable number of clusters, k.
```{r}
max_k<- 20
wcss<- rep(NA, max_k-1)

for (k in c(2:max_k)){
  k_clus<- kmeans(scaled_loc_data, centers = k, nstart = 10)
  wcss[k-1]<- k_clus$tot.withinss
}

wcss_df<- data.frame(k=c(2:max_k), wcss = wcss)
wcss_fig<- wcss_df %>% ggplot(aes(x=k, y=wcss)) + 
  geom_line() + geom_point()

print(wcss_fig)
```
Here, we see that at around k=8 the total sumsquared decreases slowly. Thus we will pick optimal k as 8 to do the clustering.
```{r}
best_k<- 8

clus_loc<- kmeans(scaled_loc_data, centers = scaled_loc_data[1:best_k,], nstart = 10)
clus_grp<- data.frame(clus_loc$cluster) %>% setNames("cluster")
merged_res<- merge(clean_loc_data, clus_grp, by ='row.names', all=TRUE)
merged_res$cluster<- as.factor(merged_res$cluster)

### center of each cluster
center_data<- merged_res %>% group_by(cluster) %>%
  summarise(num_city = n(),
            cen_mean_so2 = mean(mean_so2, na.rm=TRUE),
            cen_mean_no2 = mean(mean_no2, na.rm=TRUE),
            cen_mean_rspm = mean(mean_rspm, na.rm=TRUE),
  ) %>% data.frame()

print(center_data)
```
The above table shows the centroid for each cluster. To better visualise the difference among the clusters, we can plot the box plot of each features for the clusters:
```{r}
clus_vis_so2<- merged_res %>% ggplot(aes(x=cluster, y=mean_so2)) +
  geom_boxplot()
print(clus_vis_so2)

clus_vis_no2<- merged_res %>% ggplot(aes(x=cluster, y=mean_no2)) +
  geom_boxplot()
print(clus_vis_no2)

clus_vis_rspm<- merged_res %>% ggplot(aes(x=cluster, y=mean_rspm)) +
  geom_boxplot()
print(clus_vis_rspm)
```

From the plot, we see that cities in cluster 6, 7 and 8 have high values in all 3 features, which represents low air quality. Interestingly, these clusters have significantly high value in exactly 1 of the feature: cluster 6 is high in RSPM, cluster 7 in SO2 and cluster 8 in NO2.
For the others, cities in cluster 5 have the best air quality with lowest value in all 3 features. Cluster 2, 3, 4 have moderate value in all 3 features while cluster 1 has moderate value in SO2 and NO2 but with high RSPM.
